Unsupervised embeddings of organizations capture latent structure of scientific mobility

### 2020-03-04 paper/current/main.tex
replace f as t
replace_t_with_T,_rephrase_title_of_the_first_para

Scientific mobility drives innovation~\autocite{braunerhjelm2020labor, kaiser2018innovation, armano2017innovation}, impact~\autocite{sugimoto2017mostimpact, petersen2018multiscale, jonkers2013return, franzoni2014advantage}, collaboration~\autocite{rodrigues2016mobility}, and the diffusion of knowledge~\autocite{braunerhjelm2020labor, azoulay2011diffusion, morgan2018prestige}. 
Common models of mobility, such as the Gravity Model~\autocite{zipf1946gravity, xia2005measlesgravity, jung2008highwaygravity, barbosa2018human, curiel2018citygravity, lewer2008immigrationgravity, hong2016busgravity, truscott2012epidemicgravity, xia2005measlesgravity, simini2012universal}, describes the of mobility using \textit{distance} between places.
However, amid a globalizing scientific enterprise~\autocite{czaika2018globalisation}, geographic distance is losing its relevance for understanding mobility, whereas other abstract notions of distance~\autocite{boschma2005proximity} pose challenges; too many choices.
Accordingly, the complexity of factors driving global scientific mobility~\autocite{robinson2019mobility, vannoorden2012mobility, deville2014career, ackers2008mobility, scott2015dynamics} necessitates simplified models for reasoning about and crafting effective policy. 
Here, we demonstrate a data-driven approach of learning an effective distance that encodes the latent structure of mobility using a neural embedding model~\autocite{mikolov2013word2vec} by sourcing publications of disambiguated authors from the Web of Science database between 2008 and 2019. 
Distance between vectors in the learned embedding space explain and predict mobility between organizations better than  geographic distance or network distance. We argue that such power emerges from an underlying connection between embeddings and the gravity model.
In addition to geography, these embeddings also capture latent religious, cultural, social, and linguistic, and even prestige of organizations at multiple levels of granularity.
With this approach, we can learn scalable, dense, and meaningful vector representations of scientific mobility directly from the data and it can be generalized methodology of reasoning about all kinds of mobility.
\end{abstract}

@@

We use the trajectories of the three million name-disambiguated authors who were \textit{mobile}---having more than one affiliation---between 2008 and 2019, as evidenced by their publications indexed in the Web of Science databse. 
We adopt the Gravity Model of mobility~\autocite{zipf1946gravity, barbosa2018human}, one of the most common and broadly applied mobility models~\autocite{curiel2018citygravity, lewer2008immigrationgravity, jung2008highwaygravity, hong2016busgravity, truscott2012epidemicgravity, xia2005measlesgravity} to assess how well the embedding captures mobility. After fitting the parameters of the gravity model, we derive the \textit{expected flux} $\hat{t_{ij}}$ from the obtained gravity model, defined as,

\begin{equation}
	\label{eq:gravity_basic}
	\hat{t_{ij}} = Cm_{i}m_{j}f(r_{ij})
\end{equation}

Where the flux, or number of authors co-affiliated with two organizations during the time period $t_{ij}$, increases with the organization's sizes, $m_{i}$ and $m_{j}$, and decays as a function of their distance, $f(r_{ij})$. 
We consider two distances, the geographic distance between the organizations as a baseline, and the \textit{embedding distance} defined as the cosine distance between the organization's vector in the trained embedding space. 
We argue that the performance of a distance measure can be observed from their ability to explain expected mobility between organizations, as well as to predict flux using the gravity model. 
For geographic distance, we define $f(r_{ij})$ as the standard power-law function, and for the embedding distance, we use the exponential function. selected as the best performing for each case.

%
% Subsection: Embedding performance
%
\subsection*{Embedding performance}

@@

Let $t_{ij}$ represents the total \textbf{flux}, the number of people who moved between organizations $i$ and $j$. $M_{i}$ and $M_{j}$ represent the size of an organization. According to our gravity model on the embedding space,

\begin{equation}
    \label{eq:gravity_model}
    \frac{f_{ij}}{M_{i} M_{j}} \propto \exp(\cos(v_i,v_j))
\end{equation}

@@

In the case of scientific co-affiliation, all movements are without direction (symmetric), meaning that  $f_{ij} = f_{ji}$.
Let's denote $\mathpzc{N}_i$ as the neighborhood set of organization i, which are organizations with non-zero co-affiliation to organization i.
Then, the total flux to or from $i$ can be written as

\begin{equation}
    \mu_{i} = \sum_{k\in\mathpzc{N}_i} f_{ik} \propto M_i
\end{equation}

@@

\begin{equation}
    \label{eq:gravity_model_approx}
    \frac{f_{ij}}{\mu_i  \mu_j} \propto \exp(\cos(v_i,v_j))
\end{equation}

@@

\begin{equation}
    \label{eq:flux_conditional_prob}
    p(j|i) = \frac{f_{ij}}{\mu_i} 
\end{equation}

@@

\begin{equation}
    \label{eq:gravity_relationship}
    \frac{f_{ij}}{\mu_i  \mu_j} = \frac{\mu_i  p(j|i)}{\mu_i\mu_j} =  \frac{p(j|i)}{\mu_j} \propto \exp(\cos(v_i,v_j))
\end{equation}

>>>>

Scientific mobility drives innovation~\autocite{braunerhjelm2020labor, kaiser2018innovation, armano2017innovation}, impact~\autocite{sugimoto2017mostimpact, petersen2018multiscale, jonkers2013return, franzoni2014advantage}, collaboration~\autocite{rodrigues2016mobility}, and the diffusion of knowledge~\autocite{braunerhjelm2020labor, azoulay2011diffusion, morgan2018prestige}. 
Common models of mobility, such as the Gravity Model~\autocite{zipf1946gravity, xia2005measlesgravity, jung2008highwaygravity, barbosa2018human, curiel2018citygravity, lewer2008immigrationgravity, hong2016busgravity, truscott2012epidemicgravity, xia2005measlesgravity, simini2012universal}, describes mobility flow using \textit{distance} between places.
However, amid a globalizing scientific enterprise~\autocite{czaika2018globalisation}, geographic distance is losing its relevance for understanding mobility, whereas other abstract notions of distance~\autocite{boschma2005proximity} pose challenges; too many choices.
Accordingly, the complexity of factors driving global scientific mobility~\autocite{robinson2019mobility, vannoorden2012mobility, deville2014career, ackers2008mobility, scott2015dynamics} necessitates simplified models for reasoning about and crafting effective policy. 
Here, we demonstrate a data-driven approach of learning an effective distance that encodes the latent structure of mobility using a neural embedding model~\autocite{mikolov2013word2vec} by sourcing publications of disambiguated authors from the Web of Science database between 2008 and 2019. 
Distance between vectors in the learned embedding space explain and predict mobility between organizations better than  geographic distance or network distance. We argue that such power emerges from an underlying connection between embeddings and the gravity model.
In addition to geography, these embeddings also capture latent religious, cultural, social, and linguistic, and even prestige of organizations at multiple levels of granularity.
With this approach, we can learn scalable, dense, and meaningful vector representations of scientific mobility directly from the data and it can be generalized methodology of reasoning about all kinds of mobility.
\end{abstract}

@@

We use the trajectories of the three million name-disambiguated authors who were \textit{mobile}---having more than one affiliation---between 2008 and 2019, as evidenced by their publications indexed in the Web of Science databse. 
We adopt the Gravity Model of mobility~\autocite{zipf1946gravity, barbosa2018human}, one of the most common and broadly applied mobility models~\autocite{curiel2018citygravity, lewer2008immigrationgravity, jung2008highwaygravity, hong2016busgravity, truscott2012epidemicgravity, xia2005measlesgravity} to assess how well the embedding captures mobility. After fitting the parameters of the gravity model, we derive the \textit{expected flux} $\hat{T_{ij}}$ from the obtained gravity model, defined as,

\begin{equation}
	\label{eq:gravity_basic}
	\hat{T_{ij}} = Cm_{i}m_{j}f(r_{ij})
\end{equation}

Where the flux, or number of authors co-affiliated with two organizations during the time period $T_{ij}$, increases with the organization's sizes, $m_{i}$ and $m_{j}$, and decays as a function of their distance, $f(r_{ij})$. 
We consider two distances, the geographic distance between the organizations as a baseline, and the \textit{embedding distance} defined as the cosine distance between the organization's vector in the trained embedding space. 
We argue that the performance of a distance measure can be observed from their ability to explain expected mobility between organizations, as well as to predict flux using the gravity model. 
For geographic distance, we define $f(r_{ij})$ as the standard power-law function, and for the embedding distance, we use the exponential function. selected as the best performing for each case.

%
% Subsection: Embedding performance
%
\subsection*{Embeddings estimate the effective distance between organizations}

@@

Let $T_{ij}$ represents the total \textbf{flux}, the number of people who moved between organizations $i$ and $j$. $M_{i}$ and $M_{j}$ represent the size of an organization. According to our gravity model on the embedding space,

\begin{equation}
    \label{eq:gravity_model}
    \frac{T_{ij}}{M_{i} M_{j}} \propto \exp(\cos(v_i,v_j))
\end{equation}

@@

In the case of scientific co-affiliation, all movements are without direction (symmetric), meaning that  $T_{ij} = T_{ji}$.
Let's denote $\mathpzc{N}_i$ as the neighborhood set of organization i, which are organizations with non-zero co-affiliation to organization i.
Then, the total flux to or from $i$ can be written as

\begin{equation}
    \mu_{i} = \sum_{k\in\mathpzc{N}_i} t_{ik} \propto M_i
\end{equation}

@@

\begin{equation}
    \label{eq:gravity_model_approx}
    \frac{T_{ij}}{\mu_i  \mu_j} \propto \exp(\cos(v_i,v_j))
\end{equation}

@@

\begin{equation}
    \label{eq:flux_conditional_prob}
    p(j|i) = \frac{T_{ij}}{\mu_i} 
\end{equation}

@@

\begin{equation}
    \label{eq:gravity_relationship}
    \frac{T_{ij}}{\mu_i  \mu_j} = \frac{\mu_i  p(j|i)}{\mu_i\mu_j} =  \frac{p(j|i)}{\mu_j} \propto \exp(\cos(v_i,v_j))
\end{equation}
